---
title: "MARP Stage 1 Analysis Strategy"
author: "Don van Ravenzwaaij"
date: "December 2, 2020"
output:
  # pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

MARP Stage 1 Analysis Strategy
==============================================================

The research questions are:

1) Do religious people report higher well-being?
2) Does the relation between religiosity and well-being depend on how important people consider religion to be in their country (i.e., perceived cultural norms of religion)?

# 0. Reading in the data

First, let's read in the data:

```{r message=FALSE}
library ("BayesFactor")
```

```{r echo=FALSE}
# setwd ("/Users/donvanravenzwaaij/Dropbox/Research/UHD2 RUG/MARP")
```

```{r}
PreData = read.csv(here::here("data/MARP_data_blinded.csv"))
```

# 1. Outlier removal

For our outlier removal, we keep everything extremely simple and remove just those that failed to pass the attention check:

```{r}
Data = PreData[PreData$attention_check==1,]
```

There are no clear indications that anything else merits removal.

# 2. The IV, the CV, and the DV

We do not have any information on the psychometric properties of any of the included items in the data set. As such, we have no theoretical basis to decide which items are and are not relevant to the constructs religiosity and well-being. As such, we decided to operationalize these concepts as self-reported religiosity and self-reported well-being. Furthermore, we make no assumptions as to whether the included items are necessary and sufficient.

## The IV

The item that in our opinion reflects self-reported religiosity best is item rel_3: ``self-identification (1= religious, 2= not religious, 3=atheist)''. We note that in the data set, item codes 1, 0.5, and 0 are used, respectively. We are not interested in the distinction between not religious (coded 0.5) and atheist (coded 0), so we lump these together for further analysis.

```{r}
IV = as.factor ((Data$rel_3==1)+0)
```

## The CV

In order to answer the second research question, it is important to operationalize ``how important people consider religion to be in their country''. We have two items that potentially qualify:
- cnorm_1: importance of religious lifestyle for average person in country (1-5)
- cnorm_2: importance of belief in God/Gods for average person in country (1-5)
We have no clear indication that one is more important than the other, so we average over both.

```{r}
CV = (Data$cnorm_1+Data$cnorm_2)/2
```

## The DV

The item that in our opinion reflects self-reported well-being best is item wb_gen_1: ``quality of life general''. The item uses a 1-5 response scale, with 1 indicating low well-being and 5 indicating high well-being. With all of this defined, we can create our non-religious group (coded nRel) and our religious group (coded Rel):


```{r}
DV = Data$wb_gen_1

AnData = data.frame (IV, DV, CV)
write.csv (AnData, file = "AnData.csv")
```

# 3. Descriptives

Here are some descriptives:
```{r echo=FALSE}
ns = table (IV, CV)
Means = SDs = matrix (, dim(ns)[1]+1, dim(ns)[2]+1)
IVvals = 0:1
CVvals = seq (0, 1, 1/8)
for (i in 1:dim(ns)[1])
{
  for (j in 1:dim(ns)[2])
  {
    # Joint means and sds
    Means[i,j] = mean (DV[IV==IVvals[i]&CV==CVvals[j]])
    SDs[i,j] = sd (DV[IV==IVvals[i]&CV==CVvals[j]])
  }
  # Religion marginal means and sds
  Means[i,dim(ns)[2]+1] = mean (DV[IV==IVvals[i]])
  SDs[i,dim(ns)[2]+1] = sd (DV[IV==IVvals[i]])
}

for (j in 1:dim(ns)[2])
{
  # Cultural norm marginal means and sds
  Means[dim(ns)[1]+1,j] = mean (DV[CV==CVvals[j]])
  SDs[dim(ns)[1]+1,j] = sd (DV[CV==CVvals[j]])
}

# Grand mean and sd
Means[dim(ns)[1]+1,dim(ns)[2]+1] = mean (DV)
SDs[dim(ns)[1]+1,dim(ns)[2]+1] = sd (DV)
Means = round (Means, 2); SDs = round (SDs, 2)

# Ns per cell and in total
ns = cbind (ns, apply (ns, 1, sum))
ns = rbind (ns, apply (ns, 2, sum))
dimnames (Means) = dimnames (SDs) = dimnames (ns) = list (c(IVvals, "marginal"), c(CVvals, "marginal"))
```

```{r echo=FALSE}
knitr::kable(Means, caption = "Cell means and marginal means")
knitr::kable(SDs, caption = "Cell SDs and marginal sds")
knitr::kable(ns, caption = "Sample size per cell and across cells")
```

# 4. Analyses

Our analysis of choice is a Bayesian ANCOVA with default priors (for documentation, see https://www.rdocumentation.org/packages/BayesFactor/versions/0.9.12-4.2/topics/lmBF). For the priors, we set the $r$ parameter of the inverse gamma prior on $\sqrt2/4$ for the continuous effect (i.e., the CV), which is regarded as a "medium" prior. We set the $r$ parameter of the inverse chi square prior on $0.5$ for the fixed effect (i.e., the IV), which is regarded as a "medium" prior. Both of these settings are consistent with the JASP recommendation. In order to verify robustness, we also include analyses for $r$ parameters of $\sqrt2/2$ and $1$ for continuous and fixed effects respectively, which are considered "ultrawide" priors. We test whether there is a main effect of the IV on the DV and whether there is evidence for the IV+CV model over the null model.

<!-- This analysis is the one-sided t-test variant, don't think it needs to be included. -->
<!-- Specifically, we calculate a Jeffreys-Zellner-Siow Bayes factor, which means we put a Cauchy prior of effect size $\delta$. The Cauchy prior is centered on zero with scale parameter $1/\sqrt2$. -->
<!-- ```{r} -->
<!-- BF10 = as.vector (ttestBF (x = DV[IV==0], y = DV[IV==1], nullInterval = c(-Inf, 0),  -->
<!--                           rscale = 1/sqrt(2)))[1] -->
<!-- ``` -->

Conducting the analyses yields:
```{r}
MC = sqrt(2)/4
MF = 0.5
ModelIV = lmBF (DV~IV, data = AnData, rscaleCont = MC, rscaleFixed = MF)
BFIV = round (exp (ModelIV@bayesFactor[1]), 3)
ModelIVCV = lmBF (DV~IV+CV, data = AnData, rscaleCont = MC, rscaleFixed = MF)
BFIVCV = round (exp (ModelIVCV@bayesFactor[1]), 3)
```

The first Bayes factor indicates the data is `r BFIV` times more likely under the IV model (the $\beta$ parameter for IV religion is non-zero) compared to the null model (the $\beta$ parameter for IV religion is zero).

The second Bayes factor indicates the data is `r BFIVCV` times more likely under the IV+CV model (the $\beta$ parameters for IV religion and CV cultural context are non-zero) compared to the null model (both $\beta$ parameters are zero).

Dividing the second Bayes factor by the first gives us the gain of the CV to the model: `r round (BFIVCV/BFIV, 3)`.

Examination of the descriptives yields...

In order to verify the robustness of our results, we repeated the analyses with all combinations of one or both priors set to "ultrawide":
```{r echo=FALSE}
UC = sqrt(2)/2
UF = 1
ModelIVmu = lmBF (DV~IV, data = AnData, rscaleCont = UC, rscaleFixed = MF)
BFIVmu = round (exp (ModelIVmu@bayesFactor[1]), 3)
ModelIVCVmu = lmBF (DV~IV+CV, data = AnData, rscaleCont = UC, rscaleFixed = MF)
BFIVCVmu = round (exp (ModelIVCVmu@bayesFactor[1]), 3)
ModelIVum = lmBF (DV~IV, data = AnData, rscaleCont = MC, rscaleFixed = UF)
BFIVum = round (exp (ModelIVum@bayesFactor[1]), 3)
ModelIVCVum = lmBF (DV~IV+CV, data = AnData, rscaleCont = MC, rscaleFixed = UF)
BFIVCVum = round (exp (ModelIVCVum@bayesFactor[1]), 3)
ModelIVuu = lmBF (DV~IV, data = AnData, rscaleCont = UC, rscaleFixed = UF)
BFIVuu = round (exp (ModelIVuu@bayesFactor[1]), 3)
ModelIVCVuu = lmBF (DV~IV+CV, data = AnData, rscaleCont = UC, rscaleFixed = UF)
BFIVCVuu = round (exp (ModelIVCVuu@bayesFactor[1]), 3)
Robust = matrix (c(BFIVmu, BFIVCVmu, round (BFIVCVmu/BFIVmu, 3), 
                   BFIVum, BFIVCVum, round (BFIVCVum/BFIVum, 3),
                   BFIVuu, BFIVCVuu, round (BFIVCVuu/BFIVuu, 3)), 3, 3, 
                 dimnames = list (c("IV/null", "IV+CV/null", "(IV+CV)/IV"), c("IVmCVu", "IVuCVm", "IVuCVu")))
```

which resulted in Bayes factors as displayed below:

```{r echo=FALSE}
knitr::kable(Robust, caption = "Robustness Bayes factors for the IV vs null, the IV+CV vs null, and the IV+CV vs IV models (rows) and one or two ultawide priors for the IV or CV (columns)")
```


